---
title: "11 Strings with Stringr"
author: "Aaron Williams"
date: "7/14/2017"
output: html_document
---

```{r setup rmarkdown, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup r}
library(tidyverse)
library(stringr)
```

##### 1. In code that doesn't use `stringr`, you'll often see `paste()` and `paste0()`. What's the difference between the two functions? What `stringr` function are they equivalent to? How do the functions differ in their handling of NA?

`paste()` automatically includes a space between each character string it combines. `paste0()` does not include a space. They are ~equivalent to `str_c()` from `library(stringr)`. `paste()` and `paste0()` include NA as text. `str_c()` returns an NA for the entire string if the string contains an NA. 

##### 2. In your own words, describe the difference between the `sep` and `collapse` arguments to `str_c()`.

`sep` is a character string to insert between input vectors. Its input vector and output vector always have the same length. 

```{r sep}
length(str_c("Letter", letters, sep = ": "))
```

`collapse` is a character string to insert between input vectors and to turn the vector into a single string. `collapse` always returns a vector with length one. 

```{r collapse}
length(str_c("Letter", letters, collapse = ": "))
```

##### 3. Use `str_length()` and `str_sub()` to extract the middle character from a string. What will you do if the string has an even number of characters. 

```{r str_length}
string_middle <- function(string) {
  string_length <- str_length(string)
  
  if (string_length %% 2 == 1) {
    str_sub(string, floor((string_length + 1) / 2), ceiling((string_length) / 2))
  } 
  else if (string_length %% 2 == 0) {
    NULL
  } 
  else {"Error!"}
}

string_middle("abc")
string_middle("abcd")

```

I returned a NULL if `string_length()` is even. 

##### 4. What does `str_wrap()` do? When might you want to use it?

It implements the Knuth-Plass paragraph wrapping algorithm. It "breaks text paragraphs into lines, of total width - if it is possible - of at most given `width`.

```{r str_wrap}

graph <- "It implements the Knuth-Plass paragraph wrapping algorithm. It breaks text paragraphs into lines, of total width - if it is possible - of at most given width."

str_wrap(graph, width = 20)
```

This could be useful for formatting in html and rmarkdown. Especially for graphics and sidebars. Custom width is useful - especially in reproducible documents. 

##### 5. What does `str_trim()` do? What's the opposite of `str_trim()`?

It trims whitespace from the left, right, or both sides of a character string. It is the string version of `trimws()`.

`str_pad()` is the opposite of `str_trim()`. It adds whitespace to the left, right, or both sides of a character string. 

##### 6. Write a function that turns (e.g.) a vector `c("a", "b", "c")` into a string a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2. 

```{r str_function}





```